---
title: "Genetic Structure Analysis"
author: "Laura H Spencer"
date: "4/10/2021"
output: html_document
---
```{r, message=FALSE, warning=FALSE, results=FALSE}
list.of.packages <- c("tidyverse", "reshape2", "plotly", "vcfR", "SNPRelate", "network", "janitor", "vcfR") #add new libraries here 
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

# Load all libraries 
lapply(list.of.packages, FUN = function(X) {
  do.call("require", list(X)) 
})
```

## Load sample info and create dataframes for each stage 

```{r}
load(file="../raw-data/sample.info")

sample.info.adult <- sample.info %>%
          select(sample, stage, population, pCO2.parent, temp.parent) %>% filter(stage=="adult") %>% 
 mutate(pop_code = as.factor(paste(population, pCO2.parent, sep="-"))) %>% droplevels()

sample.info.larvae <- sample.info %>%
          select(sample, stage, population, pCO2.parent) %>% filter(stage=="larvae" & sample!=571) %>% 
 mutate(pop_code = as.factor(paste(population, pCO2.parent, sep="-")),
        sample=str_pad(sample, width=3, side="left", pad="0")) %>% droplevels()

sample.info.juvenile <- sample.info %>%
          select(sample, stage, population, pCO2.parent) %>% filter(stage=="juvenile") %>% 
 mutate(pop_code = as.factor(paste(population, pCO2.parent, sep="-")),
        sample=str_pad(sample, width=3, side="left", pad="0")) %>% droplevels()
```

# ADULTS, COMPARISON AMONG PCO2 TREATMENTS WITHIN POPULATIONS 

## Use gatk to hard filter variants for depth & quality

```{bash}
/Applications/bioinformatics/gatk-4.1.9.0/gatk VariantFiltration \
-R ../references/Olurida_v081_concat_rehead.fa \
-V ../qc-processing/gatk/Olurida_QuantSeq2020_genotypes-adult.vcf.gz \
-O ../qc-processing/gatk/Olurida_QuantSeq2020_genotypes-adult-filtered.vcf.gz \
--filter-name "FS" \
--filter "FS > 60.0" \
--filter-name "QUAL30" \
--filter "QUAL < 30.0" \
--filter-name "SOR3" \
--filter "SOR > 3.0" \
--filter-name "DP50" \
--filter "DP < 50" \
--filter-name "DP450" \
--filter "DP > 450"
```

## Select only SNPs that pass filtering

```{bash}
/Applications/bioinformatics/gatk-4.1.9.0/gatk SelectVariants \
-R ../references/Olurida_v081_concat_rehead.fa \
-V ../qc-processing/gatk/Olurida_QuantSeq2020_genotypes-adult-filtered.vcf.gz \
--exclude-filtered TRUE \
--select-type-to-include SNP \
-O ../qc-processing/gatk/Olurida_QuantSeq2020_genotypes-adult-filtered-true.vcf.gz
```
## Use SNPRelate to filter vcf files. Remove: 
- Select biallelic loci only 
- Loci with minor allele frequency < 5%
- Select loci that are present across all samples (missing.rate=0) 

```{r}
#snpgdsClose(genofile.adult)
vcf.fn.adult <- "../qc-processing/gatk/Olurida_QuantSeq2020_genotypes-adult-filtered-true.vcf.gz"
snpgdsVCF2GDS(vcf.fn.adult, "../qc-processing/gatk/genotypes-adult.gds", method="biallelic.only")
snpgdsSummary("../qc-processing/gatk/genotypes-adult.gds")
```

## Open gds and check out genofile contents 

```{r}
(genofile.adult <- snpgdsOpen("../qc-processing/gatk/genotypes-adult.gds"))
```

## Prune/Clean SNPs 
- Remove those in linkage-disequilibrium
- Remove those with Minor Allele Frequency <1%
- Remove those that are missing from any sample 

```{r}
snpset.adult <- snpgdsLDpruning(genofile.adult, maf=.1, missing.rate=0, autosome.only=FALSE)
snpset.adult.id <- unlist(unname(snpset.adult))
```

# For ADULT genotype analysis:  ~2,950 SNPs remain after all filtering steps 

## Adult PCA  

```{r}
# PCA 
pca.adult <- snpgdsPCA(genofile.adult, snp.id=snpset.adult.id, num.thread=2, autosome.only=FALSE)
pc.percent.adult <- pca.adult$varprop*100
head(round(pc.percent.adult, 2))
tab.adult <- data.frame(sample.id = pca.adult$sample.id,
    EV1 = pca.adult$eigenvect[,1],    # the first eigenvector
    EV2 = pca.adult$eigenvect[,2],    # the second eigenvector
    EV3 = pca.adult$eigenvect[,3],    # the third eigenvector
    EV4 = pca.adult$eigenvect[,4],    # the fourth eigenvector
    EV5 = pca.adult$eigenvect[,5],    # the fifth eigenvector
    stringsAsFactors = FALSE)
tab.adult.annot <- left_join(tab.adult, sample.info.adult, by=c("sample.id"="sample")) %>% droplevels()
group.adult <- as.factor(tab.adult.annot$pop_code)

#ggplotly( 
    ggplot(tab.adult.annot, aes(x=EV2, y=EV1, col=population, shape=pCO2.parent)) + 
         geom_point(size=4) + theme_minimal() + ggtitle("Adult Genetic Structure, PC1xPC2")#)

ggplotly( 
    ggplot(tab.adult.annot, aes(x=EV3, y=EV1, col=population, shape=pCO2.parent)) + 
         geom_point(size=4) + theme_minimal() + ggtitle("Adult Genetic Structure, PC1xPC3"))

ggplotly( 
    ggplot(tab.adult.annot, aes(x=EV4, y=EV1, col=population, shape=pCO2.parent)) + 
         geom_point(size=4) + theme_minimal() + ggtitle("Adult Genetic Structure, PC1xPC4"))

ggplotly( 
    ggplot(tab.adult.annot, aes(x=EV5, y=EV1, col=population, shape=pCO2.parent)) + 
         geom_point(size=4) + theme_minimal() + ggtitle("Adult Genetic Structure, PC1xPC5"))
```

## Adult cluster analysis 

```{r}
# To perform cluster analysis on the matrix of genome-wide IBS pairwise distances, 
# and determine the groups by a permutation score. 
# Result= only 1 group (?)
ibs.hc.adult <- snpgdsHCluster(snpgdsIBS(genofile.adult, snp.id=snpset.adult.id, num.thread=2, autosome.only=FALSE))
rv.adult <- snpgdsCutTree(ibs.hc.adult)
plot(rv.adult$dendrogram, leaflab="none", main="HapMap Phase II")

# Determine groups of individuals by population information
rv2.adult <- snpgdsCutTree(ibs.hc.adult, samp.group=as.factor(tab.adult.annot$pop_code))
plot(rv2.adult$dendrogram, leaflab="none", main="HapMap Phase II")
legend("bottom", legend=levels(group.adult), col=1:nlevels(group.adult), pch=19, ncol=4, cex=0.6, pt.cex=1)
```

## Adult Allele frequencies 

```{r}
AF.adult.FB.high <- snpgdsSNPRateFreq(genofile.adult, snp.id=snpset.adult.id, 
                        sample.id=subset(tab.adult.annot, population=="Fidalgo Bay" & pCO2.parent=="High")$sample.id,
                        with.id = TRUE)
AF.adult.FB.amb <- snpgdsSNPRateFreq(genofile.adult, snp.id=snpset.adult.id, 
                        sample.id=subset(tab.adult.annot, population=="Fidalgo Bay" & pCO2.parent=="Ambient")$sample.id,
                        with.id = TRUE)
AF.adult.DB.high <- snpgdsSNPRateFreq(genofile.adult, snp.id=snpset.adult.id, 
                        sample.id=subset(tab.adult.annot, population=="Dabob Bay" & pCO2.parent=="High")$sample.id,
                        with.id = TRUE)
AF.adult.DB.amb <- snpgdsSNPRateFreq(genofile.adult, snp.id=snpset.adult.id, 
                        sample.id=subset(tab.adult.annot, population=="Dabob Bay" & pCO2.parent=="Ambient")$sample.id,
                        with.id = TRUE)
AF.adult.OB1.high <- snpgdsSNPRateFreq(genofile.adult, snp.id=snpset.adult.id, 
                        sample.id=subset(tab.adult.annot, population=="Oyster Bay C1" & pCO2.parent=="High")$sample.id,
                        with.id = TRUE)
AF.adult.OB1.amb <- snpgdsSNPRateFreq(genofile.adult, snp.id=snpset.adult.id, 
                        sample.id=subset(tab.adult.annot, population=="Oyster Bay C1" & pCO2.parent=="Ambient")$sample.id,
                        with.id = TRUE)

AF.adults <- rbind(do.call(cbind.data.frame, AF.adult.FB.high[c("snp.id", "AlleleFreq")]) %>% mutate(population=as.factor("FB"), stage=as.factor("adult"), pCO2=as.factor("high")), 
do.call(cbind.data.frame, AF.adult.FB.amb[c("snp.id", "AlleleFreq")]) %>% mutate(population=as.factor("FB"), stage=as.factor("adult"), pCO2=as.factor("ambient")), 
do.call(cbind.data.frame, AF.adult.DB.high[c("snp.id", "AlleleFreq")]) %>% mutate(population=as.factor("DB"), stage=as.factor("adult"), pCO2=as.factor("high")), 
do.call(cbind.data.frame, AF.adult.DB.amb[c("snp.id", "AlleleFreq")]) %>% mutate(population=as.factor("DB"), stage=as.factor("adult"), pCO2=as.factor("ambient")),
do.call(cbind.data.frame, AF.adult.OB1.high[c("snp.id", "AlleleFreq")]) %>% mutate(population=as.factor("OB1"), stage=as.factor("adult"), pCO2=as.factor("high")), 
do.call(cbind.data.frame, AF.adult.OB1.amb[c("snp.id", "AlleleFreq")]) %>% mutate(population=as.factor("OB1"), stage=as.factor("adult"), pCO2=as.factor("ambient")))

ggplot(AF.adults, aes(x=pCO2, y=AlleleFreq, fill=population, alpha=pCO2)) + geom_violin(trim = F) + facet_wrap(~population) + theme_minimal() +scale_alpha_discrete(range = c(0.35, 0.9)) + 
  stat_summary(fun.y=mean, geom="point", shape=8, size=4, color="black", fill="black") + xlab("pCO2 exposure") + ylab("Allele Frequency") +
  ggtitle(paste("SNP Allele Frequency, Adults by population & pCO2 treatment\n", "(No. loci = ", length(snpset.adult.id), ")", sep="")) +
    geom_text(data=AF.adults %>% group_by(population, pCO2) %>% summarize(mean=mean(AlleleFreq)), aes(x=pCO2, y=mean+.1, label=round(mean, 2)))
```
## Adult Fst Calculations 

### Note: snpgdsFst() returns 5 things:
  - sample IDs 
  - SNP ids (i.e. indices)
  - "Fst" = weighted Fst estimate 
  - "MeanFst" = average of Fst estimates across all SNPs 
  - "FstSNP" = A vector of Fst for each SNP 

```{r}
# within FB (comparing pCO2 exposure): 
Fst.adult.FB <- snpgdsFst(genofile.adult, snp.id=snpset.adult.id, 
                        sample.id=subset(tab.adult.annot, population=="Fidalgo Bay")$sample.id, 
               population=as.factor(subset(tab.adult.annot, population=="Fidalgo Bay")$pCO2.parent),
    method="W&C84", autosome.only=FALSE, with.id = TRUE)
paste("Weir and Cockerham mean Fst estimate, Fidalgo Bay Ambient vs. High: ", round(Fst.adult.FB$MeanFst, 5))
hist(Fst.adult.FB$FstSNP, breaks = 75, main="Fst distribution\nFidalgo Bay Ambient vs. High, Adults", col="dodgerblue3")

# within DB (comparing pCO2 exposure): 
Fst.adult.DB <- snpgdsFst(genofile.adult, snp.id=snpset.adult.id, 
                        sample.id=subset(tab.adult.annot, population=="Dabob Bay")$sample.id, 
               population=as.factor(subset(tab.adult.annot, population=="Dabob Bay")$pCO2.parent),
    method="W&C84", autosome.only=FALSE, with.id = TRUE)
paste("Weir and Cockerham mean Fst estimate, Dabob Bay Ambient vs. High: ", round(Fst.adult.DB$MeanFst, 5))
hist(Fst.adult.DB$FstSNP, breaks = 100, main="Fst distribution\nDabob Bay Ambient vs. High, Adults", col="darkseagreen3")

# within OB1 (comparing pCO2 exposure): 
Fst.adult.OB1 <- snpgdsFst(genofile.adult, snp.id=snpset.adult.id, 
                         sample.id=subset(tab.adult.annot, population=="Oyster Bay C1")$sample.id, 
               population=as.factor(subset(tab.adult.annot, population=="Oyster Bay C1")$pCO2.parent),
    method="W&C84", autosome.only=FALSE, with.id = TRUE)
paste("Weir and Cockerham mean Fst estimate, Oyster Bay C1 Ambient vs. High: ", round(Fst.adult.OB1$MeanFst, 5))
hist(Fst.adult.OB1$FstSNP, breaks = 100, main="Fst distribution\nOyster Bay C1 Ambient vs. High, Adults", col="salmon2")

Fst.adults <- rbind(do.call(cbind.data.frame, Fst.adult.FB[c("snp.id", "FstSNP")]) %>% mutate(population=as.factor("FB"), stage=as.factor("adult")), 
do.call(cbind.data.frame, Fst.adult.DB[c("snp.id", "FstSNP")]) %>% mutate(population=as.factor("DB"), stage=as.factor("adult")),
do.call(cbind.data.frame, Fst.adult.OB1[c("snp.id", "FstSNP")]) %>% mutate(population=as.factor("OB1"), stage=as.factor("adult"))) 

ggplot(Fst.adults, aes(x=population, y=FstSNP, fill=population)) + geom_violin(trim = F) + 
  stat_summary(fun.y=mean, geom="point", shape=8, size=3.5, color="black", fill="black") +
  ggtitle(paste("SNP Fst, Adults by pCO2 treatment\n", "(No. loci = ", length(snpset.adult.id), ")", sep="")) +
    geom_text(data=Fst.adults %>% group_by(population) %>% summarize(mean=mean(FstSNP)), aes(x=population, y=.65, label=round(mean, 3)))
```
## Adult Identity-by-state 

```{r}
lbls <- paste("PC", 1:4, "\n", format(pc.percent.adult[1:4], digits=2), "%", sep="")
pairs(pca.adult$eigenvect[,1:4], col=as.color(tab.adult.annot$pCO2.parent), labels=lbls, main="colors = pCO2 only")
pairs(pca.adult$eigenvect[,1:4], col=as.color(tab.adult.annot$pop_code), labels=lbls, main="colors = pop & pCO2")
```

```{r}
ibs.adult <- snpgdsIBS(genofile.adult, snp.id=snpset.adult.id, num.thread=2, autosome.only=FALSE)
pop.idx.adult <- order(tab.adult.annot$pop_code)

#par(xpd=TRUE)
#heatmap(ibs.adult$ibs, col=terrain.colors(16), labCol=TRUE, 
#        RowSideColors=c("black","red","green","blue")[tab.adult.annot$pop_code])
#legend(0,-.05, legend=levels(tab.adult.annot$pop_code), pch="o", col=1:nlevels(tab.adult.annot$pop_code), 
#       text.col=1:nlevels(tab.adult.annot$pop_code), cex=.75, ncol=4)

#Perform multidimensional scaling analysis on the matrix of genome-wide IBS pairwise distances:
loc.adult <- cmdscale(1 - ibs.adult$ibs, k = 2)
x.adult <- loc.adult[, 1]; y.adult <- loc.adult[, 2]

plot(x.adult, y.adult, col=group.adult, xlab = "", ylab = "",
    main = "Multidimensional Scaling Analysis (IBS)", cex=2, pch=16)
text(x.adult, y.adult+.01, labels=tab.adult.annot$sample.id, cex=0.7, font=2, col=as.integer(group.adult))
legend("bottom", legend=levels(group.adult), pch="o", text.col=1:nlevels(group.adult))
```

# LARVAE, COMPARISON AMONG PCO2 TREATMENTS WITHIN POPULATIONS 

## Use gatk to hard filter variants for depth & quality

```{bash}
/Applications/bioinformatics/gatk-4.1.9.0/gatk VariantFiltration \
-R ../references/Olurida_v081_concat_rehead.fa \
-V ../qc-processing/gatk/Olurida_QuantSeq2020_genotypes-larvae.vcf.gz \
-O ../qc-processing/gatk/Olurida_QuantSeq2020_genotypes-larvae-filtered.vcf.gz \
--filter-name "FS" \
--filter "FS > 60.0" \
--filter-name "QUAL30" \
--filter "QUAL < 30.0" \
--filter-name "SOR3" \
--filter "SOR > 3.0" \
--filter-name "DP50" \
--filter "DP < 50" \
--filter-name "DP450" \
--filter "DP > 450"
```

## Select only SNPs that pass filtering

```{bash}
/Applications/bioinformatics/gatk-4.1.9.0/gatk SelectVariants \
-R ../references/Olurida_v081_concat_rehead.fa \
-V ../qc-processing/gatk/Olurida_QuantSeq2020_genotypes-larvae-filtered.vcf.gz \
--exclude-filtered TRUE \
--select-type-to-include SNP \
-O ../qc-processing/gatk/Olurida_QuantSeq2020_genotypes-larvae-filtered-true.vcf.gz
```

## Use SNPRelate to filter vcf files. Remove: 
- Select biallelic loci only 
- Loci with minor allele frequency < 0.3 (?)
- Select loci that are present across all samples (missing.rate=0) 

```{r}
#snpgdsClose(genofile.larvae)
vcf.fn.larvae <- "../qc-processing/gatk/Olurida_QuantSeq2020_genotypes-larvae-filtered-true.vcf.gz"
snpgdsVCF2GDS(vcf.fn.larvae, "../qc-processing/gatk/genotypes-larvae.gds", method="biallelic.only")
snpgdsSummary("../qc-processing/gatk/genotypes-larvae.gds")
```

## Open gds and check out genofile contents 

```{r}
(genofile.larvae <- snpgdsOpen("../qc-processing/gatk/genotypes-larvae.gds"))
```

## Prune/Clean SNPs 
- Remove those in linkage-disequilibrium
- Remove those with Minor Allele Frequency <1%
- Remove those that are missing from any sample 

```{r}
snpset.larvae <- snpgdsLDpruning(genofile.larvae, maf=.10, missing.rate=0.10, autosome.only=FALSE)
snpset.larvae.id <- unlist(unname(snpset.larvae))
```

# For LARVAL analysis, ~360 SNPs remain after all filtering steps 

## Larval PCA 

```{r}
# PCA 
pca.larvae <- snpgdsPCA(genofile.larvae, snp.id=snpset.larvae.id, num.thread=2, autosome.only=FALSE)
pc.percent.larvae <- pca.larvae$varprop*100
head(round(pc.percent.larvae, 2))
tab.larvae <- data.frame(sample.id = pca.larvae$sample.id,
    EV1 = pca.larvae$eigenvect[,1],    # the first eigenvector
    EV2 = pca.larvae$eigenvect[,2],    # the second eigenvector
    EV3 = pca.larvae$eigenvect[,3],    # the third eigenvector
    EV4 = pca.larvae$eigenvect[,4],    # the fourth eigenvector
    EV5 = pca.larvae$eigenvect[,5],    # the fourth eigenvector
    stringsAsFactors = FALSE)
tab.larvae.annot <- left_join(tab.larvae, sample.info.larvae, by=c("sample.id"="sample")) %>% droplevels()
group.larvae <- as.factor(tab.larvae.annot$pop_code)

lbls <- paste("PC", 1:5, "\n", format(pc.percent.larvae[1:5], digits=2), "%", sep="")
pairs(pca.larvae$eigenvect[,1:5], col=as.color(tab.larvae.annot$pCO2.parent), labels=lbls, main="colors = pCO2 only")
pairs(pca.larvae$eigenvect[,1:5], col=as.color(tab.larvae.annot$pop_code), labels=lbls, main="colors = pop & pCO2")
```

Interactive PCA biplots 

```{r}
# PC1 x PC2
# uncomment out the ggplotly lines to interact
#ggplotly( 
    ggplot(tab.larvae.annot, aes(x=EV2, y=EV1, col=population, shape=pCO2.parent)) + 
         geom_point(size=4) + theme_minimal() + ggtitle("PC1 x PC2\nLarvae Genetic Structure (pooled batches)")#)
#PC1 x PC3
# uncomment out the ggplotly lines to interact
ggplotly( 
    ggplot(tab.larvae.annot, aes(x=EV3, y=EV1, col=population, shape=pCO2.parent)) + 
         geom_point(size=4) + theme_minimal() + ggtitle("PC1 x PC3\nLarvae Genetic Structure (pooled batches)"))
# PC1 x PC4
ggplotly( 
    ggplot(tab.larvae.annot, aes(x=EV4, y=EV2, col=population, shape=pCO2.parent)) + 
         geom_point(size=4) + theme_minimal() + ggtitle("PC1 x PC4\nLarvae Gene Expression (pooled batches)"))
# PC1 x PC5
ggplotly( 
    ggplot(tab.larvae.annot, aes(x=EV5, y=EV1, col=population, shape=pCO2.parent)) + 
         geom_point(size=4) + theme_minimal() + ggtitle("PC1 x PC5\nLarvae Gene Expression (pooled batches)"))
```

## Larval cluster analysis 

```{r}
# To perform cluster analysis on the matrix of genome-wide IBS pairwise distances, 
# and determine the groups by a permutation score. 
ibs.hc.larvae <- snpgdsHCluster(snpgdsIBS(genofile.larvae, num.thread=2,  snp.id=snpset.larvae.id, autosome.only=FALSE, 
                                         sample.id=setdiff(tab.larvae.annot$sample.id, "044")))
rv.larvae <- snpgdsCutTree(ibs.hc.larvae)
plot(rv.larvae$dendrogram, leaflab="none", main="HapMap Phase II")

# Determine groups of individuals by population information
group.larvae <- as.factor(tab.larvae.annot$pop_code)
rv2.larvae <- snpgdsCutTree(ibs.hc.larvae, samp.group=as.factor(subset(tab.larvae.annot, sample.id!="044")$pop_code))
plot(rv2.larvae$dendrogram, leaflab="none", main="HapMap Phase II")
legend("top", legend=levels(group.larvae), col=1:nlevels(group.larvae), pch=19, ncol=4, cex=0.6, pt.cex=1)
```
## Larval Allele frequencies 

```{r}
AF.larvae.FB.high <- snpgdsSNPRateFreq(genofile.larvae, snp.id=snpset.larvae.id, 
                        sample.id=subset(tab.larvae.annot, population=="Fidalgo Bay" & pCO2.parent=="High")$sample.id,
                        with.id = TRUE)
AF.larvae.FB.amb <- snpgdsSNPRateFreq(genofile.larvae, snp.id=snpset.larvae.id, 
                        sample.id=subset(tab.larvae.annot, population=="Fidalgo Bay" & pCO2.parent=="Ambient")$sample.id,
                        with.id = TRUE)
AF.larvae.DB.high <- snpgdsSNPRateFreq(genofile.larvae, snp.id=snpset.larvae.id, 
                        sample.id=subset(tab.larvae.annot, population=="Dabob Bay" & pCO2.parent=="High")$sample.id,
                        with.id = TRUE)
AF.larvae.DB.amb <- snpgdsSNPRateFreq(genofile.larvae, snp.id=snpset.larvae.id, 
                        sample.id=subset(tab.larvae.annot, population=="Dabob Bay" & pCO2.parent=="Ambient")$sample.id,
                        with.id = TRUE)
AF.larvae.OB1.high <- snpgdsSNPRateFreq(genofile.larvae, snp.id=snpset.larvae.id, 
                        sample.id=subset(tab.larvae.annot, population=="Oyster Bay C1" & pCO2.parent=="High")$sample.id,
                        with.id = TRUE)
AF.larvae.OB1.amb <- snpgdsSNPRateFreq(genofile.larvae, snp.id=snpset.larvae.id, 
                        sample.id=subset(tab.larvae.annot, population=="Oyster Bay C1" & pCO2.parent=="Ambient")$sample.id,
                        with.id = TRUE)
AF.larvae.OB2.high <- snpgdsSNPRateFreq(genofile.larvae, snp.id=snpset.larvae.id, 
                        sample.id=subset(tab.larvae.annot, population=="Oyster Bay C2" & pCO2.parent=="High")$sample.id,
                        with.id = TRUE)
AF.larvae.OB2.amb <- snpgdsSNPRateFreq(genofile.larvae, snp.id=snpset.larvae.id, 
                        sample.id=subset(tab.larvae.annot, population=="Oyster Bay C2" & pCO2.parent=="Ambient")$sample.id,
                        with.id = TRUE)

AF.larvae <- rbind(do.call(cbind.data.frame, AF.larvae.FB.high[c("snp.id", "AlleleFreq")]) %>% mutate(population=as.factor("FB"), stage=as.factor("larvae"), pCO2=as.factor("high")), 
do.call(cbind.data.frame, AF.larvae.FB.amb[c("snp.id", "AlleleFreq")]) %>% mutate(population=as.factor("FB"), stage=as.factor("larvae"), pCO2=as.factor("ambient")), 
do.call(cbind.data.frame, AF.larvae.DB.high[c("snp.id", "AlleleFreq")]) %>% mutate(population=as.factor("DB"), stage=as.factor("larvae"), pCO2=as.factor("high")), 
do.call(cbind.data.frame, AF.larvae.DB.amb[c("snp.id", "AlleleFreq")]) %>% mutate(population=as.factor("DB"), stage=as.factor("larvae"), pCO2=as.factor("ambient")),
do.call(cbind.data.frame, AF.larvae.OB1.high[c("snp.id", "AlleleFreq")]) %>% mutate(population=as.factor("OB1"), stage=as.factor("larvae"), pCO2=as.factor("high")), 
do.call(cbind.data.frame, AF.larvae.OB1.amb[c("snp.id", "AlleleFreq")]) %>% mutate(population=as.factor("OB1"), stage=as.factor("larvae"), pCO2=as.factor("ambient")),
do.call(cbind.data.frame, AF.larvae.OB2.high[c("snp.id", "AlleleFreq")]) %>% mutate(population=as.factor("OB2"), stage=as.factor("larvae"), pCO2=as.factor("high")), 
do.call(cbind.data.frame, AF.larvae.OB2.amb[c("snp.id", "AlleleFreq")]) %>% mutate(population=as.factor("OB2"), stage=as.factor("larvae"), pCO2=as.factor("ambient")))

ggplot(AF.larvae, aes(x=pCO2, y=AlleleFreq, fill=population, alpha=pCO2)) + geom_violin(trim = F) + facet_wrap(~population, ncol=4) + theme_minimal() +scale_alpha_discrete(range = c(0.35, 0.9)) + 
  stat_summary(fun.y=mean, geom="point", shape=8, size=4, color="black", fill="black") + xlab("pCO2 exposure") + ylab("Allele Frequency") +
  ggtitle(paste("SNP Allele Frequency, larvaes by population & pCO2 treatment\n", "(No. loci = ", length(snpset.larvae.id), ")", sep="")) +
    geom_text(data=AF.larvae %>% group_by(population, pCO2) %>% summarize(mean=mean(AlleleFreq)), aes(x=pCO2, y=mean+.1, label=round(mean, 2))) + theme(legend.position = "none")
```

## Larval Fst Calculations 

```{r}
# within FB (comparing pCO2 exposure): 
Fst.larvae.FB <- snpgdsFst(genofile.larvae, snp.id=snpset.larvae.id, 
                        sample.id=subset(tab.larvae.annot, population=="Fidalgo Bay")$sample.id, 
               population=as.factor(subset(tab.larvae.annot, population=="Fidalgo Bay")$pCO2.parent),
    method="W&C84", autosome.only=FALSE, with.id = TRUE)
paste("Weir and Cockerham mean Fst estimate, Fidalgo Bay Ambient vs. High: ", round(Fst.larvae.FB$MeanFst, 5))
hist(Fst.larvae.FB$FstSNP, breaks = 75, main="Fst distribution\nFidalgo Bay Ambient vs. High, Larvae", col="dodgerblue3", xlim = c(-.15, 0.8))

# within DB (comparing pCO2 exposure): 
Fst.larvae.DB <- snpgdsFst(genofile.larvae, snp.id=snpset.larvae.id, 
                        sample.id=subset(tab.larvae.annot, population=="Dabob Bay")$sample.id, 
               population=as.factor(subset(tab.larvae.annot, population=="Dabob Bay")$pCO2.parent),
    method="W&C84", autosome.only=FALSE, with.id = TRUE)
paste("Weir and Cockerham mean Fst estimate, Dabob Bay Ambient vs. High: ", round(Fst.larvae.DB$MeanFst, 5))
hist(Fst.larvae.DB$FstSNP, breaks = 100, main="Fst distribution\nDabob Bay Ambient vs. High, Larvae", col="darkseagreen3", xlim = c(-.2, .8))

# within OB1 (comparing pCO2 exposure): 
Fst.larvae.OB1 <- snpgdsFst(genofile.larvae, snp.id=snpset.larvae.id, 
                         sample.id=subset(tab.larvae.annot, population=="Oyster Bay C1")$sample.id, 
               population=as.factor(subset(tab.larvae.annot, population=="Oyster Bay C1")$pCO2.parent),
    method="W&C84", autosome.only=FALSE, with.id = TRUE)
paste("Weir and Cockerham mean Fst estimate, Oyster Bay C1 Ambient vs. High: ", round(Fst.larvae.OB1$MeanFst, 5))
hist(Fst.larvae.OB1$FstSNP, breaks = 100, main="Fst distribution\nOyster Bay C1 Ambient vs. High, Larvae", col="salmon2", xlim = c(-.2, 0.8))

Fst.larvae <- rbind(do.call(cbind.data.frame, Fst.larvae.FB[c("snp.id", "FstSNP")]) %>% mutate(population=as.factor("FB"), stage=as.factor("larvae")), 
do.call(cbind.data.frame, Fst.larvae.DB[c("snp.id", "FstSNP")]) %>% mutate(population=as.factor("DB"), stage=as.factor("larvae")),
do.call(cbind.data.frame, Fst.larvae.OB1[c("snp.id", "FstSNP")]) %>% mutate(population=as.factor("OB1"), stage=as.factor("larvae"))) 

ggplot(Fst.larvae, aes(x=population, y=FstSNP, fill=population)) + geom_violin(trim = F) + 
  stat_summary(fun.y=mean, geom="point", shape=8, size=3.5, color="black", fill="black") +
  ggtitle(paste("SNP Fst, Larvae by pCO2 treatment\n", "(No. loci = ", length(snpset.larvae.id), ")", sep="")) +
    geom_text(data=Fst.larvae %>% group_by(population) %>% summarize(mean=mean(FstSNP)), aes(x=population, y=.78, label=round(mean, 3))) +
  theme_minimal()
```

Larval Identify-by-state 

```{r}
ibs.larvae <- snpgdsIBS(genofile.larvae, num.thread=2, snp.id=snpset.larvae.id, autosome.only=FALSE, sample.id=setdiff(tab.larvae.annot$sample.id, "044"))
pop.idx.larvae <- order(tab.larvae.annot$pop_code)

#perform multidimensional scaling analysis on the matrix of genome-wide IBS pairwise distances:
loc.larvae <- cmdscale(1 - ibs.larvae$ibs, k = 2)
x.larvae <- loc.larvae[, 1]; y.larvae <- loc.larvae[, 2]
a <- as.factor(subset(tab.larvae.annot, sample.id!="044")$population)
b <- as.numeric(as.factor(subset(tab.larvae.annot, sample.id!="044")$pCO2.parent))

par(xpd=TRUE)
plot(x.larvae, y.larvae, col=group.larvae, xlab = "", ylab = "",
    main = "Multidimensional Scaling Analysis (IBS)", cex=2, pch=16)
#text(x.larvae, y.larvae+.01, labels=tab.larvae.annot$sample.id, cex=0.7, font=2, col=as.integer(group.larvae))
legend("bottomright", inset=c(-.065,0), legend=levels(group.larvae), cex=.6, pch="o", text.col=1:nlevels(group.larvae))

# Option to plot color coded by pop, symbol by pCO2  
#plot(x.larvae, y.larvae, col=a, pch=b, xlab = "", ylab = "",
#    main = "Multidimensional Scaling Analysis (IBS)", cex=1)
```

# JUVENILES, COMPARISON AMONG PCO2 TREATMENTS WITHIN POPULATIONS 

## Use gatk to hard filter variants for depth & quality

```{bash}
/Applications/bioinformatics/gatk-4.1.9.0/gatk VariantFiltration \
-R ../references/Olurida_v081_concat_rehead.fa \
-V ../qc-processing/gatk/Olurida_QuantSeq2020_genotypes-juvenile.vcf.gz \
-O ../qc-processing/gatk/Olurida_QuantSeq2020_genotypes-juvenile-filtered.vcf.gz \
--filter-name "FS" \
--filter "FS > 60.0" \
--filter-name "QUAL30" \
--filter "QUAL < 30.0" \
--filter-name "SOR3" \
--filter "SOR > 3.0" \
--filter-name "DP50" \
--filter "DP < 50" \
--filter-name "DP450" \
--filter "DP > 450"
```

## Select only SNPs that pass filtering

```{bash}
/Applications/bioinformatics/gatk-4.1.9.0/gatk SelectVariants \
-R ../references/Olurida_v081_concat_rehead.fa \
-V ../qc-processing/gatk/Olurida_QuantSeq2020_genotypes-juvenile-filtered.vcf.gz \
--exclude-filtered TRUE \
--select-type-to-include SNP \
-O ../qc-processing/gatk/Olurida_QuantSeq2020_genotypes-juvenile-filtered-true.vcf.gz
```

## Use SNPRelate to filter vcf files. Remove: 
- Select biallelic loci only 
- Loci with minor allele frequency < 0.3 (?)
- Select loci that are present across all samples (missing.rate=0) 

```{r}
#snpgdsClose(genofile.juvenile)
vcf.fn.juvenile <- "../qc-processing/gatk/Olurida_QuantSeq2020_genotypes-juvenile-filtered-true.vcf.gz"
snpgdsVCF2GDS(vcf.fn.juvenile, "../qc-processing/gatk/genotypes-juvenile.gds", method="biallelic.only")
snpgdsSummary("../qc-processing/gatk/genotypes-juvenile.gds")
```

## Open gds and check out genofile contents 

```{r}
(genofile.juvenile <- snpgdsOpen("../qc-processing/gatk/genotypes-juvenile.gds"))
```

## Prune/Clean SNPs 
- Remove those in linkage-disequilibrium
- Remove those with Minor Allele Frequency <1%
- Remove those that are missing from any sample 

```{r}
snpset.juvenile <- snpgdsLDpruning(genofile.juvenile, maf=.1, missing.rate=0, autosome.only=FALSE)
snpset.juvenile.id <- unlist(unname(snpset.juvenile))
```

# For JUVENILE analysis, ~2,600 SNPs remain after all filtering steps 

## Juvenile PCA 

```{r}
# PCA 
pca.juvenile <- snpgdsPCA(genofile.juvenile, snp.id=snpset.juvenile.id, num.thread=2, autosome.only=FALSE)
pc.percent.juvenile <- pca.juvenile$varprop*100
head(round(pc.percent.juvenile, 2))
tab.juvenile <- data.frame(sample.id = pca.juvenile$sample.id,
    EV1 = pca.juvenile$eigenvect[,1],    # the first eigenvector
    EV2 = pca.juvenile$eigenvect[,2],    # the second eigenvector
    EV3 = pca.juvenile$eigenvect[,3],    # the third eigenvector
    EV4 = pca.juvenile$eigenvect[,4],    # the fourth eigenvector
    EV5 = pca.juvenile$eigenvect[,5],    # the fourth eigenvector
    stringsAsFactors = FALSE)
tab.juvenile.annot <- left_join(tab.juvenile, sample.info.juvenile, by=c("sample.id"="sample")) %>% droplevels()
group.juvenile <- as.factor(tab.juvenile.annot$pop_code)

lbls <- paste("PC", 1:5, "\n", format(pc.percent.juvenile[1:5], digits=2), "%", sep="")
pairs(pca.juvenile$eigenvect[,1:5], col=as.color(tab.juvenile.annot$pCO2.parent), labels=lbls, main="colors = pCO2 only")
pairs(pca.juvenile$eigenvect[,1:5], col=as.color(tab.juvenile.annot$pop_code), labels=lbls, main="colors = pop & pCO2")
```

Interactive PCAs 

```{r}
# PC1 x PC2
# uncomment out the ggplotly lines to interact
ggplotly( 
    ggplot(tab.juvenile.annot, aes(x=EV2, y=EV1, col=population, shape=pCO2.parent)) + 
         geom_point(size=4) + theme_minimal() + ggtitle("PC1 x PC2\njuvenile Genetic Structure (pooled batches)"))
#PC1 x PC3
# uncomment out the ggplotly lines to interact
ggplotly( 
    ggplot(tab.juvenile.annot, aes(x=EV3, y=EV1, col=population, shape=pCO2.parent)) + 
         geom_point(size=4) + theme_minimal() + ggtitle("PC1 x PC3\njuvenile Genetic Structure (pooled batches)"))
#PC2 x PC3
# uncomment out the ggplotly lines to interact
ggplotly( 
    ggplot(tab.juvenile.annot, aes(x=EV3, y=EV2, col=population, shape=pCO2.parent)) + 
         geom_point(size=4) + theme_minimal() + ggtitle("PC2 x PC3\njuvenile Genetic Structure (pooled batches)"))
```

## Juvenile cluster analysis 

```{r}
# To perform cluster analysis on the matrix of genome-wide IBS pairwise distances, 
# and determine the groups by a permutation score. 
ibs.hc.juvenile <- snpgdsHCluster(snpgdsIBS(genofile.juvenile, num.thread=2,  snp.id=snpset.juvenile.id, autosome.only=FALSE, 
                                         sample.id=setdiff(tab.juvenile.annot$sample.id, "044")))
rv.juvenile <- snpgdsCutTree(ibs.hc.juvenile)
plot(rv.juvenile$dendrogram, leaflab="none", main="HapMap Phase II")

# Determine groups of individuals by population information
rv2.juvenile <- snpgdsCutTree(ibs.hc.juvenile, samp.group=as.factor(subset(tab.juvenile.annot, sample.id!="044")$pop_code))
plot(rv2.juvenile$dendrogram, leaflab="none", main="HapMap Phase II")
legend("top", legend=levels(group.juvenile), col=1:nlevels(group.juvenile), pch=19, ncol=4, cex=0.6, pt.cex=1)
```
## Juvenile Allele frequencies 

```{r}
AF.juvenile.FB.high <- snpgdsSNPRateFreq(genofile.juvenile, snp.id=snpset.juvenile.id, 
                        sample.id=subset(tab.juvenile.annot, population=="Fidalgo Bay" & pCO2.parent=="High")$sample.id,
                        with.id = TRUE)
AF.juvenile.FB.amb <- snpgdsSNPRateFreq(genofile.juvenile, snp.id=snpset.juvenile.id, 
                        sample.id=subset(tab.juvenile.annot, population=="Fidalgo Bay" & pCO2.parent=="Ambient")$sample.id,
                        with.id = TRUE)
AF.juvenile.DB.high <- snpgdsSNPRateFreq(genofile.juvenile, snp.id=snpset.juvenile.id, 
                        sample.id=subset(tab.juvenile.annot, population=="Dabob Bay" & pCO2.parent=="High")$sample.id,
                        with.id = TRUE)
AF.juvenile.DB.amb <- snpgdsSNPRateFreq(genofile.juvenile, snp.id=snpset.juvenile.id, 
                        sample.id=subset(tab.juvenile.annot, population=="Dabob Bay" & pCO2.parent=="Ambient")$sample.id,
                        with.id = TRUE)

AF.juveniles <- rbind(do.call(cbind.data.frame, AF.juvenile.FB.high[c("snp.id", "AlleleFreq")]) %>% mutate(population=as.factor("FB"), stage=as.factor("juvenile"), pCO2=as.factor("high")), 
do.call(cbind.data.frame, AF.juvenile.FB.amb[c("snp.id", "AlleleFreq")]) %>% mutate(population=as.factor("FB"), stage=as.factor("juvenile"), pCO2=as.factor("ambient")), 
do.call(cbind.data.frame, AF.juvenile.DB.high[c("snp.id", "AlleleFreq")]) %>% mutate(population=as.factor("DB"), stage=as.factor("juvenile"), pCO2=as.factor("high")), 
do.call(cbind.data.frame, AF.juvenile.DB.amb[c("snp.id", "AlleleFreq")]) %>% mutate(population=as.factor("DB"), stage=as.factor("juvenile"), pCO2=as.factor("ambient")))

ggplot(AF.juveniles, aes(x=pCO2, y=AlleleFreq, fill=population, alpha=pCO2)) + geom_violin(trim = F) + facet_wrap(~population) + theme_minimal() +scale_alpha_discrete(range = c(0.35, 0.9)) + 
  stat_summary(fun.y=mean, geom="point", shape=8, size=4, color="black", fill="black") + xlab("pCO2 exposure") + ylab("Allele Frequency") +
  ggtitle(paste("SNP Allele Frequency, juveniles by population & pCO2 treatment\n", "(No. loci = ", length(snpset.juvenile.id), ")", sep="")) +
    geom_text(data=AF.juveniles %>% group_by(population, pCO2) %>% summarize(mean=mean(AlleleFreq)), aes(x=pCO2, y=mean+.1, label=round(mean, 2)))
```

## Juvenile Fst Calculations 

```{r}
# within FB (comparing pCO2 exposure): 
Fst.juvenile.FB <- snpgdsFst(genofile.juvenile, snp.id=snpset.juvenile.id, 
                        sample.id=subset(tab.juvenile.annot, population=="Fidalgo Bay")$sample.id, 
               population=as.factor(subset(tab.juvenile.annot, population=="Fidalgo Bay")$pCO2.parent),
    method="W&C84", autosome.only=FALSE, with.id = TRUE)
paste("Weir and Cockerham mean Fst estimate, Fidalgo Bay Ambient vs. High: ", round(Fst.juvenile.FB$MeanFst, 5))
hist(Fst.juvenile.FB$FstSNP, breaks = 75, main="Fst distribution\nFidalgo Bay Ambient vs. High, juvenile", col="dodgerblue3", xlim = c(-.15, 0.8))

# within DB (comparing pCO2 exposure): 
Fst.juvenile.DB <- snpgdsFst(genofile.juvenile, snp.id=snpset.juvenile.id, 
                        sample.id=subset(tab.juvenile.annot, population=="Dabob Bay")$sample.id, 
               population=as.factor(subset(tab.juvenile.annot, population=="Dabob Bay")$pCO2.parent),
    method="W&C84", autosome.only=FALSE, with.id = TRUE)
paste("Weir and Cockerham mean Fst estimate, Dabob Bay Ambient vs. High: ", round(Fst.juvenile.DB$MeanFst, 5))
hist(Fst.juvenile.DB$FstSNP, breaks = 100, main="Fst distribution\nDabob Bay Ambient vs. High, juvenile", col="darkseagreen3", xlim = c(-.2, .8))

Fst.juvenile <- rbind(do.call(cbind.data.frame, Fst.juvenile.FB[c("snp.id", "FstSNP")]) %>% mutate(population=as.factor("FB"), stage=as.factor("juvenile")), 
do.call(cbind.data.frame, Fst.juvenile.DB[c("snp.id", "FstSNP")]) %>% mutate(population=as.factor("DB"), stage=as.factor("juvenile"))) 

ggplot(Fst.juvenile, aes(x=population, y=FstSNP, fill=population)) + geom_violin(trim = F) + 
  stat_summary(fun.y=mean, geom="point", shape=8, size=3.5, color="black", fill="black") +
  ggtitle(paste("SNP Fst, juvenile by pCO2 treatment\n", "(No. loci = ", length(snpset.juvenile.id), ")", sep="")) +
    geom_text(data=Fst.juvenile %>% group_by(population) %>% summarize(mean=mean(FstSNP)), aes(x=population, y=1.1, label=round(mean, 3))) +
  theme_minimal()
```

## Juvenile Identify-by-state 

```{r}
ibs.juvenile <- snpgdsIBS(genofile.juvenile, num.thread=2, snp.id=snpset.juvenile.id, autosome.only=FALSE, sample.id=setdiff(tab.juvenile.annot$sample.id, "044"))
pop.idx.juvenile <- order(tab.juvenile.annot$pop_code)

#perform multidimensional scaling analysis on the matrix of genome-wide IBS pairwise distances:
loc.juvenile <- cmdscale(1 - ibs.juvenile$ibs, k = 2)
x.juvenile <- loc.juvenile[, 1]; y.juvenile <- loc.juvenile[, 2]
a <- as.factor(subset(tab.juvenile.annot, sample.id!="044")$population)
b <- as.numeric(as.factor(subset(tab.juvenile.annot, sample.id!="044")$pCO2.parent))

plot(x.juvenile, y.juvenile, col=a, pch=b, xlab = "", ylab = "",
    main = "Multidimensional Scaling Analysis (IBS)", cex=1)
#text(x.juvenile, y.juvenile+.012, labels=tab.juvenile.annot$sample.id, cex=0.7, font=2, col=as.integer(group.juvenile))
#legend("topleft", legend=levels(group.juvenile), pch="o", cex=0.8) #text.col=1:nlevels(group.juvenile), 
```
# Merge all Allele Frequency and Fst calculations together and plot 

```{r}
Fst.all <- rbind(Fst.adults, Fst.larvae, Fst.juvenile)
Fst.means <- Fst.all %>% select(-snp.id) %>% group_by(population, stage) %>% summarize(mean=mean(FstSNP), sd=sd(FstSNP), n=n())

AF.all <- rbind(AF.adults, AF.larvae, AF.juveniles)
AF.means <- AF.all %>% select(-snp.id) %>% group_by(population, stage, pCO2) %>% summarize(mean=mean(AlleleFreq), sd=sd(AlleleFreq), n=n())

ggplot(AF.all, aes(x=stage, y=AlleleFreq, fill=pCO2, alpha=stage)) + geom_violin(trim = T) + theme_minimal()  +   
  ggtitle("Allele Frequency\nBy population, stage, &  adult pCO2 treatment") + xlab("") + ylab("Allele Frequency")  + facet_wrap(~population) +scale_alpha_discrete(range = c(0.85, 0.15)) #+
#  g0eom_point(data=AF.all %>% group_by(population, stage, pCO2) %>% summarize(mean=mean(AlleleFreq)), 
#             aes(x=stage:pCO2, y=mean, label=round(mean, 3)), cex=2.25, inherit.aes=T, position = position_dodge(width=1))

ggplot(Fst.all, aes(x=stage, y=FstSNP, fill=stage)) + geom_violin(trim = T, alpha=0.4) + theme_minimal() + facet_wrap(~ population) +   
    stat_summary(fun.y=mean, geom="point", aes(group=stage), shape=8, size=2, color="black", fill="black") +
  ggtitle("Per-SNP Fst among pCO2 treatments\nBy population and stage") + xlab("") + ylab("Per-Locus Fst") +
  geom_text(data=Fst.means, aes(group=stage, y=-.33, label=round(mean, 3)), cex=3.25)

ggplot(Fst.all %>% filter(population==c("FB","DB")), aes(x=FstSNP, fill=stage)) + geom_density(alpha=0.4) + theme_minimal()  +   
  ggtitle("Per-SNP Fst among adult pCO2 treatments (high vs. ambient)\nBy population & stage") + xlab("") + ylab("Per-Locus Fst")  + facet_wrap(~population)
```
# More analysis via vcfR - compare allelic richness and heterozygosity by population, stage and parental pCO2 exposure 

First filter the vcf files for each stage (adult, larvae, adult) to remove loci with tons if missing data. Do this via vcftools. 

I used [this tutorial](http://www.ddocent.com/filtering/) as an example. 
  - call vcftools
  - feed it a vcf file after the --vcf flag
  - Set --max-missing 0.90 to keep only loci with 10% missing genotypes across all individuals
  - The --recode flag tells the program to write a new vcf file with the filters
  - The --recode-INFO-all keeps all the INFO flags from the old vcf file in the new one. 
  - Lastly, --out designates the name of the output
  
```{bash}
# adults 
vcftools --gzvcf "../qc-processing/gatk/Olurida_QuantSeq2020_genotypes-adult-filtered-true.vcf.gz" \
--max-missing 0.85 --recode --recode-INFO-all --out \
"../qc-processing/gatk/Olurida_QuantSeq2020_genotypes-adult-4vcfR"
```
```{bash}
# larvae 
vcftools --gzvcf "../qc-processing/gatk/Olurida_QuantSeq2020_genotypes-larvae-filtered-true.vcf.gz" \
--max-missing 0.85 --recode --recode-INFO-all --out \
"../qc-processing/gatk/Olurida_QuantSeq2020_genotypes-larvae-4vcfR"
```

```{bash}
# juveniles 
vcftools --gzvcf "../qc-processing/gatk/Olurida_QuantSeq2020_genotypes-juvenile-filtered-true.vcf.gz" \
--max-missing 0.85 --recode --recode-INFO-all --out \
"../qc-processing/gatk/Olurida_QuantSeq2020_genotypes-juvenile-4vcfR"
```

# Adults 

```{r}
require(vcfR)
# Load the data
vcf.adult <- read.vcfR("../qc-processing/gatk/Olurida_QuantSeq2020_genotypes-adult-4vcfR.recode.vcf", verbose = FALSE )

# Confirm order of samples in vcf matches order of samples in the "sample.info" dataframe 
colnames(extract.gt(vcf.adult)) == sample.info.adult$sample #yes, all are TRUE

# Calculate differentiation / diversity stats for all population + pCO2 groups 
myDiff.adult <- genetic_diff(vcf.adult, pops = sample.info.adult$pop_code, method = 'nei')
paste("No. SNPs included in this vcfR analysis, adults ", nrow(myDiff.adult), sep="")

# Summarize diversity indices (mean)
knitr::kable(round(colMeans(myDiff.adult[,-1:-2], na.rm = TRUE), digits = 3))

# Violin plots of Heterozygosity 
dpf.adult <- melt(myDiff.adult[,c(3:8)], varnames=c('Index', 'Sample'), value.name = 'Depth', na.rm=TRUE) %>%
  mutate(population=substring(variable, 4)) %>% separate(population, into = c("population", "pCO2"), sep = "-") %>%
  mutate(population=as.factor(population), pCO2=as.factor(pCO2))

ggplot(dpf.adult, aes(x=population:pCO2, y=Depth)) + geom_violin(aes(fill=population, alpha=pCO2), adjust = 1.2) + xlab("") + ylab("") + theme_bw() +
    theme(axis.text.x = element_text(size=7, angle=50, hjust=1)) + ggtitle(paste("Heterozygosity, Adults \nNo. SNPs =", nrow(myDiff.adult))) +  
  scale_alpha_discrete(range = c(0.35, 0.9)) +
        stat_summary(fun.y=mean, geom="point", shape=15, size=3, color="black", fill="black") +
    geom_text(data=dpf.adult %>% group_by(population, pCO2) %>% summarize(mean=mean(Depth)), aes(x=population:pCO2, y=-0.05, label=round(mean, 2)), size=3) + 
  scale_alpha_discrete(guide="none")

# Boxplots of allelic richness 
dpf.n.adult <- melt(myDiff.adult[,c(10:15)], varnames=c('Index', 'Sample'), value.name = 'Depth', na.rm=TRUE) %>%
  mutate(population=substring(variable, 3)) %>% separate(population, into = c("population", "pCO2"), sep = "-") %>%
  mutate(population=as.factor(population), pCO2=as.factor(pCO2))

ggplot(dpf.n.adult, aes(x=population:pCO2, y=Depth)) + geom_violin(aes(fill=population, alpha=pCO2), adjust = 1.2) + xlab("") + ylab("") + theme_bw() +
    theme(axis.text.x = element_text(size=7, angle=50, hjust=1)) + ggtitle(paste("Allelic Richness, Adults \nNo. SNPs =", nrow(myDiff.adult))) +  
  scale_alpha_discrete(range = c(0.35, 0.9)) +
        stat_summary(fun.y=mean, geom="point", shape=15, size=3, color="black", fill="black") +
    geom_text(data=dpf.n.adult %>% group_by(population, pCO2) %>% summarize(mean=mean(Depth)), aes(x=population:pCO2, y=18.5, label=round(mean, 2)), size=3) + 
  scale_alpha_discrete(guide="none")
```

# Larvae 

```{r}
# Load the data
vcf.larvae <- read.vcfR("../qc-processing/gatk/Olurida_QuantSeq2020_genotypes-larvae-4vcfR.recode.vcf", verbose = FALSE )

# Write script to order sample.info dataframe to have same order as vcf  
colnames(extract.gt(vcf.larvae)) == sample.info.larvae[match(colnames(extract.gt(vcf.larvae)), sample.info.larvae$sample),]$sample #yes, all are TRUE

# Calculate differentiation / diversity stats for all population + pCO2 groups 
myDiff.larvae <- genetic_diff(vcf.larvae, pops = sample.info.larvae[match(colnames(extract.gt(vcf.larvae)), sample.info.larvae$sample),]$pop_code, method = 'nei')
paste("No. SNPs included in this vcfR analysis, juveniles ", nrow(myDiff.larvae), sep="")

# Summarize diversity indices (mean)
knitr::kable(round(colMeans(myDiff.larvae[,-1:-2], na.rm = TRUE), digits = 3))

# Violin plots of Heterozygosity
dpf.larvae <- melt(myDiff.larvae[,c(3:10)], varnames=c('Index', 'Sample'), value.name = 'Depth', na.rm=TRUE) %>%
  mutate(population=substring(variable, 4)) %>% separate(population, into = c("population", "pCO2"), sep = "-") %>%
  mutate(population=as.factor(population), pCO2=as.factor(pCO2))

ggplot(dpf.larvae, aes(x=population:pCO2, y=Depth)) + geom_violin(aes(fill=population, alpha=pCO2), adjust = 1.2) + xlab("") + ylab("") + theme_bw() +
    theme(axis.text.x = element_text(size=7, angle=50, hjust=1)) + ggtitle(paste("Heterozygosity, Larvae \nNo. SNPs =", nrow(myDiff.larvae))) + 
  scale_alpha_discrete(range = c(0.35, 0.9)) +
        stat_summary(fun.y=mean, geom="point", shape=15, size=3, color="black", fill="black") +
    geom_text(data=dpf.larvae %>% group_by(population, pCO2) %>% summarize(mean=mean(Depth)), aes(x=population:pCO2, y=-0.05, label=round(mean, 2)), size=3)  + 
  scale_alpha_discrete(guide="none")

# Boxplots of allelic richness 
dpf.n.larvae <- melt(myDiff.larvae[,c(12:19)], varnames=c('Index', 'Sample'), value.name = 'Depth', na.rm=TRUE) %>%
  mutate(population=substring(variable, 3)) %>% separate(population, into = c("population", "pCO2"), sep = "-") %>%
  mutate(population=as.factor(population), pCO2=as.factor(pCO2))

ggplot(dpf.n.larvae, aes(x=population:pCO2, y=Depth)) + geom_violin(aes(fill=population, alpha=pCO2), adjust = 1.2) + xlab("") + ylab("") + theme_bw() +
    theme(axis.text.x = element_text(size=7, angle=50, hjust=1)) + ggtitle(paste("Allelic Richness, Larvae \nNo. SNPs =", nrow(myDiff.larvae))) + 
  scale_alpha_discrete(range = c(0.35, 0.9)) +
        stat_summary(fun.y=mean, geom="point", shape=15, size=3, color="black", fill="black") +
    geom_text(data=dpf.n.larvae %>% group_by(population, pCO2) %>% summarize(mean=mean(Depth)), aes(x=population:pCO2, y=3, label=round(mean, 2)), size=3)  + 
  scale_alpha_discrete(guide="none")
```
# 1-yr old "juveniles"

```{r}
# Load the data
vcf.juvenile <- read.vcfR("../qc-processing/gatk/Olurida_QuantSeq2020_genotypes-juvenile-4vcfR.recode.vcf", verbose = FALSE )

# Confirm order of samples in vcf matches order of samples in the "sample.info" dataframe 
colnames(extract.gt(vcf.juvenile)) == sample.info.juvenile$sample #yes, all are TRUE

# Calculate differentiation / diversity stats for all population + pCO2 groups 
myDiff.juvenile <- genetic_diff(vcf.juvenile, pops = sample.info.juvenile$pop_code, method = 'nei')
paste("No. SNPs included in this vcfR analysis, juveniles ", nrow(myDiff.juvenile), sep="")

# Summarize diversity indices (mean)
knitr::kable(round(colMeans(myDiff.juvenile[,-1:-2], na.rm = TRUE), digits = 3))

# Violin plots of Heterozygosity
dpf.juvenile <- melt(myDiff.juvenile[,c(3:6)], varnames=c('Index', 'Sample'), value.name = 'Depth', na.rm=TRUE) %>%
  mutate(population=substring(variable, 4)) %>% separate(population, into = c("population", "pCO2"), sep = "-") %>%
  mutate(population=as.factor(population), pCO2=as.factor(pCO2))

ggplot(dpf.juvenile, aes(x=population:pCO2, y=Depth)) + geom_violin(aes(fill=population, alpha=pCO2), adjust = 1.2) + xlab("") + ylab("") + theme_bw() +
    theme(axis.text.x = element_text(size=7, angle=50, hjust=1)) + ggtitle(paste("Heterozygosity, Juveniles \nNo. SNPs =", nrow(myDiff.juvenile))) + 
  scale_alpha_discrete(range = c(0.35, 0.9)) +
        stat_summary(fun.y=mean, geom="point", shape=15, size=3, color="black", fill="black") +
    geom_text(data=dpf.juvenile %>% group_by(population, pCO2) %>% summarize(mean=mean(Depth)), aes(x=population:pCO2, y=-0.05, label=round(mean, 2)), size=3)  + 
  scale_alpha_discrete(guide="none")

# Boxplots of allelic richness 
dpf.n.juvenile <- melt(myDiff.juvenile[,c(8:11)], varnames=c('Index', 'Sample'), value.name = 'Depth', na.rm=TRUE) %>%
  mutate(population=substring(variable, 3)) %>% separate(population, into = c("population", "pCO2"), sep = "-") %>%
  mutate(population=as.factor(population), pCO2=as.factor(pCO2))

ggplot(dpf.n.juvenile, aes(x=population:pCO2, y=Depth)) + geom_violin(aes(fill=population, alpha=pCO2), adjust = 1.2) + xlab("") + ylab("") + theme_bw() +
    theme(axis.text.x = element_text(size=7, angle=50, hjust=1)) + ggtitle(paste("Allelic Richness, Juveniles \nNo. SNPs =", nrow(myDiff.juvenile))) + 
  scale_alpha_discrete(range = c(0.35, 0.9)) +
        stat_summary(fun.y=mean, geom="point", shape=15, size=3, color="black", fill="black") +
    geom_text(data=dpf.n.juvenile %>% group_by(population, pCO2) %>% summarize(mean=mean(Depth)), aes(x=population:pCO2, y=8.2, label=round(mean, 2)), size=3)  + 
  scale_alpha_discrete(guide="none")
```


```{r}
myDiff.AL <- inner_join(myDiff.adult, myDiff.larvae, by=c("CHROM", "POS"))  #7,202 loci common among all
paste("No. SNPs shared among adults and larvae: ", nrow(myDiff.AL), sep="")

# Compare adults vs. larvae 
knitr::kable(round(colMeans(inner_join(myDiff.adult, myDiff.larvae, by=c("CHROM", "POS"))[,-1:-2], na.rm = TRUE), digits = 3))

# Heterozygosity 
dpf.AL <- 
  melt(inner_join(myDiff.adult, myDiff.larvae, by=c("CHROM", "POS")) [,c(3:8,20:27)], 
               varnames=c('Index', 'Sample'), value.name = 'Depth', na.rm=TRUE) %>%
    mutate(variable=gsub("\\.x", ".adult", variable)) %>%
    mutate(variable=gsub("\\.y", ".larvae", variable)) %>%
  mutate(population=substring(variable, 4)) %>% separate(population, into = c("population", "pCO2", "stage"), sep = '[-.]') %>%
  mutate_each_(funs(factor(.)),c("population", "pCO2", "stage"))
  
ggplot(subset(dpf.AL, population!="Oyster Bay C2")%>%droplevels(), aes(x=population:stage:pCO2, y=Depth)) + 
  geom_violin(aes(fill=population, alpha=pCO2, col=stage), adjust = 1.2) + xlab("") + ylab("") + theme_bw() +
    theme(axis.text.x = element_text(size=7, angle=50, hjust=1)) + ggtitle(paste("Heterozygosity, Adults & Larvae\nNo. shared SNPs =", nrow(myDiff.AL))) + 
  scale_alpha_discrete(range = c(0.35, 0.9)) +
        stat_summary(fun.y=mean, geom="point", shape=15, size=3, color="black", fill="black") +
    geom_text(data=dpf.AL %>% group_by(population, stage, pCO2) %>% summarize(mean=mean(Depth)), aes(x=population:stage:pCO2, y=-0.05, label=round(mean, 2)), size=3) +
  scale_colour_manual(values=c("black", "dodgerblue3")) + 
  scale_alpha_discrete(guide="none")

# Allelic Richness 
dpf.n.AL <- melt(inner_join(myDiff.adult, myDiff.larvae, by=c("CHROM", "POS")) [,c(10:15,29:36)], 
               varnames=c('Index', 'Sample'), value.name = 'Depth', na.rm=TRUE) %>%
    mutate(variable=gsub("\\.x", ".adult", variable)) %>%
    mutate(variable=gsub("\\.y", ".larvae", variable)) %>%
  mutate(population=substring(variable, 3)) %>% separate(population, into = c("population", "pCO2", "stage"), sep = '[-.]') %>%
  mutate_each_(funs(factor(.)),c("population", "pCO2", "stage"))

ggplot(subset(dpf.n.AL, population!="Oyster Bay C2")%>%droplevels(), aes(x=population:stage:pCO2, y=Depth)) + 
  geom_violin(aes(fill=population, alpha=pCO2, col=stage), width=1) + xlab("") + ylab("") + theme_bw() +
    theme(axis.text.x = element_text(size=7, angle=50, hjust=1.3)) + ggtitle(paste("Allelic Richness, Adults & Larvae\nNo. shared SNPs =", nrow(myDiff.AL))) + 
  scale_alpha_discrete(range = c(0.35, 0.9)) +
        stat_summary(fun.y=mean, geom="point", shape=15, size=3, color="black", fill="black") +
    geom_text(data=dpf.n.AL %>% group_by(population, stage, pCO2) %>% summarize(mean=mean(Depth)), aes(x=population:stage:pCO2, y=2, label=round(mean, 2)), size=3) +
  scale_colour_manual(values=c("black", "dodgerblue3")) + 
  scale_alpha_discrete(guide="none")
```

# Compare adults and 1-yr old offspring 

```{r}
myDiff.AJ <- inner_join(myDiff.adult, myDiff.juvenile, by=c("CHROM", "POS"))
paste("No. SNPs shared among adults and juveniles: ", nrow(myDiff.AJ), sep="")

# Compare adults vs. larvae 
knitr::kable(round(colMeans(myDiff.AJ[,-1:-2], na.rm = TRUE), digits = 3))

# Heterozygosity 
dpf.AJ <- melt(select(myDiff.AJ,contains("Hs")), 
               varnames=c('Index', 'Sample'), value.name = 'Depth', na.rm=TRUE) %>%
    mutate(variable=gsub("\\.x", ".adult", variable)) %>%
    mutate(variable=gsub("\\.y", ".juvenile", variable)) %>%
  mutate(population=substring(variable, 4)) %>% separate(population, into = c("population", "pCO2", "stage"), sep = '[-.]') %>%
  mutate_each_(funs(factor(.)),c("population", "pCO2", "stage"))
  
ggplot(filter(dpf.AJ, !grepl('Oyster', population))%>%droplevels(), aes(x=population:stage:pCO2, y=Depth)) + 
  geom_violin(aes(fill=population, alpha=pCO2, col=stage), adjust = 1.2) + xlab("") + ylab("") + theme_bw() +
    theme(axis.text.x = element_text(size=7, angle=50, hjust=1)) + ggtitle(paste("Heterozygosity, Adults & Juveniles\nNo. shared SNPs =", nrow(myDiff.AJ))) + 
  scale_alpha_discrete(range = c(0.35, 0.9)) +
        stat_summary(fun.y=mean, geom="point", shape=15, size=3, color="black", fill="black") +
    geom_text(data=filter(dpf.AJ, !grepl('Oyster', population))%>%droplevels() %>% group_by(population, stage, pCO2) %>% summarize(mean=mean(Depth)), aes(x=population:stage:pCO2, y=-0.05, label=round(mean, 2)), size=3) +
  scale_colour_manual(values=c("black", "darkred")) + 
  scale_alpha_discrete(guide="none")

# Allelic Richness 
dpf.n.AJ <- melt(select(myDiff.AJ,contains("n_")), 
               varnames=c('Index', 'Sample'), value.name = 'Depth', na.rm=TRUE) %>%
    mutate(variable=gsub("\\.x", ".adult", variable)) %>%
    mutate(variable=gsub("\\.y", ".juvenile", variable)) %>%
  mutate(population=substring(variable, 3)) %>% separate(population, into = c("population", "pCO2", "stage"), sep = '[-.]') %>%
  mutate_each_(funs(factor(.)),c("population", "pCO2", "stage"))

ggplot(filter(dpf.n.AJ, !grepl('Oyster', population))%>%droplevels(), aes(x=population:stage:pCO2, y=Depth)) + 
  geom_violin(aes(fill=population, alpha=pCO2, col=stage), adjust = 1.2) + xlab("") + ylab("") + theme_bw() +
    theme(axis.text.x = element_text(size=7, angle=50, hjust=1)) + ggtitle(paste("Heterozygosity, Adults & Juveniles\nNo. shared SNPs =", nrow(myDiff.AJ))) + 
  scale_alpha_discrete(range = c(0.35, 0.9)) +
        stat_summary(fun.y=mean, geom="point", shape=15, size=3, color="black", fill="black") +
    geom_text(data=filter(dpf.n.AJ, !grepl('Oyster', population))%>%droplevels() %>% group_by(population, stage, pCO2) %>% summarize(mean=mean(Depth)), aes(x=population:stage:pCO2, y=-0.05, label=round(mean, 2)), size=3) +
  scale_colour_manual(values=c("black", "darkred")) + 
  scale_alpha_discrete(guide="none")
```

# Compare adults, larvae, and 1-yr old offspring 

```{r}
myDiff.ALJ <- inner_join(inner_join(myDiff.adult, myDiff.larvae, by=c("CHROM", "POS")), myDiff.juvenile,  by=c("CHROM", "POS")) 
paste("No. SNPs shared among adults, larvae, and juveniles: ", nrow(myDiff.ALJ), sep="")

# Compare adults vs. larvae 
knitr::kable(round(colMeans(myDiff.ALJ[,-1:-2], na.rm = TRUE), digits = 3))

# Heterozygosity 
dpf.ALJ <- melt(select(myDiff.ALJ,contains("Hs")) %>%  rename_at(vars(`Hs_Dabob Bay-Ambient`:`Hs_Fidalgo Bay-High`), paste0, ".z"), 
               varnames=c('Index', 'Sample'), value.name = 'Depth', na.rm=TRUE) %>%
    mutate(variable=gsub("\\.x", ".adult", variable)) %>%
    mutate(variable=gsub("\\.y", ".larvae", variable)) %>%
    mutate(variable=gsub("\\.z", ".juvenile", variable)) %>%
    mutate(population=substring(variable, 4)) %>% separate(population, into = c("population", "pCO2", "stage"), sep = '[-.]') %>%
    mutate_each_(funs(factor(.)),c("population", "pCO2", "stage")) %>%
  mutate(stage=fct_relevel(stage, c("adult", "larvae", "juvenile")))

ggplot(subset(dpf.ALJ, population!="Oyster Bay C2")%>% droplevels(), aes(x=population:stage:pCO2, y=Depth)) + 
  geom_violin(aes(fill=stage, alpha=pCO2), adjust = 1.2) + xlab("Parental pCO2 exposure") + ylab("") + theme_bw() +
    theme(axis.text.x = element_text(size=7, angle=50, hjust=1)) + ggtitle(paste("Heterozygosity: Adults, Larvae, and Juveniles\nNo. shared SNPs =", nrow(myDiff.ALJ))) + 
  scale_alpha_discrete(range = c(0.35, 0.9)) +
        stat_summary(fun.y=mean, geom="point", shape=15, size=3, color="black", fill="black") +
    geom_text(data=subset(dpf.ALJ, population!="Oyster Bay C2")%>%droplevels() %>% group_by(population, stage, pCO2) %>% summarize(mean=mean(Depth)), 
              aes(x=population:stage:pCO2, y=-0.05, label=round(mean, 2)), size=3) +
  scale_alpha_discrete(guide="none") + 
  annotate("text", x = 3, y = .6, label = "Dabob Bay", size=3.5) +
  annotate("text", x = 9, y = .6, label = "Fidalgo Bay", size=3.5) +
  annotate("text", x = 14.5, y = .6, label = "Oyster Bay", size=3.5) +
  geom_vline(xintercept=6.5) + geom_vline(xintercept=12.5) + 
  scale_x_discrete(labels=rep(c("Ambient", "High"), times=8)) 

# Allelic Richness 
dpf.n.ALJ <- melt(select(myDiff.ALJ,contains("n_")) %>%  rename_at(vars(`n_Dabob Bay-Ambient`:`n_Fidalgo Bay-High`), paste0, ".z"), 
               varnames=c('Index', 'Sample'), value.name = 'Depth', na.rm=TRUE) %>%
    mutate(variable=gsub("\\.x", ".adult", variable)) %>%
    mutate(variable=gsub("\\.y", ".larvae", variable)) %>%
    mutate(variable=gsub("\\.z", ".juvenile", variable)) %>%
    mutate(population=substring(variable, 3)) %>% separate(population, into = c("population", "pCO2", "stage"), sep = '[-.]') %>%
    mutate_each_(funs(factor(.)),c("population", "pCO2", "stage")) %>%
  mutate(stage=fct_relevel(stage, c("adult", "larvae", "juvenile")))

ggplot(subset(dpf.n.ALJ, population!="Oyster Bay C2")%>%droplevels(), aes(x=population:stage:pCO2, y=Depth)) + 
  geom_violin(aes(fill=stage, alpha=pCO2),width=1.3) + xlab("Parental pCO2 exposure") + ylab("") + theme_bw() +
    theme(axis.text.x = element_text(size=7, angle=50, hjust=1)) + ggtitle(paste("Allelic Richness: Adults, Larvae, and Juveniles\nNo. shared SNPs =", nrow(myDiff.ALJ))) + 
  scale_alpha_discrete(range = c(0.35, 0.9)) +
        #stat_summary(fun.y=mean, geom="point", shape=15, size=2, color="black", fill="black") +
    geom_text(data=subset(dpf.n.ALJ, population!="Oyster Bay C2")%>%droplevels() %>% group_by(population, stage, pCO2) %>% summarize(mean=mean(Depth)), 
              aes(x=population:stage:pCO2, y=3, label=round(mean, 2)), size=2.5) +
  #scale_colour_manual(values=c("black", "dodgerblue3", "darkred")) + 
  scale_alpha_discrete(guide="none") + geom_vline(xintercept=6.5) + geom_vline(xintercept=12.5) + 
  annotate("text", x = 3, y = 30, label = "Dabob Bay", size=3.5) +
  annotate("text", x = 9, y = 30, label = "Fidalgo Bay", size=3.5) +
  annotate("text", x = 13.5, y = 30, label = "Oyster\nBay", size=3.5) +
  scale_x_discrete(labels=rep(c("Ambient", "High"), times=8))
```

# Identify siblings using Colony 

I will use Colony to identify siblings in my samples. The Colony software is only available via the command line for Mac users (it's typically used via GUI on Windows). To get the most current version of the program I actually had to email the developer. He send me a link to download it. I'm running Colony Version 2.0.6.6 (June 30, 2020), which is saved on my computer and in this project's repo (references/colony2.mac.20200904).

Prepare files for Colony sibship analysis
Colony identifies siblings, which I would like to do with my juveniles. Here I prepare files for Colony.

Mac suggested that I only keep SNPs with very high minor allele frequencies. Here I use SNPRelate to export SNPs with MAF >=2%. 

## Adult Sibship analysis via Colony 

```{r}
snpset.adult.4Colony <- snpgdsLDpruning(genofile.adult, maf=.2, missing.rate=0, autosome.only=FALSE)
snpset.adult.id.4Colony <- unlist(unname(snpset.adult.4Colony))
snpgdsCreateGenoSet(src.fn="../qc-processing/gatk/genotypes-adult.gds",
                    dest.fn="../qc-processing/colony/genotypes-adult-4Colony.gds",
                    snp.id=snpset.adult.id.4Colony)
```

## Recode genotypes for Colony 

- **SNPRelate codes alleles this way:** "There are four possible values stored in the variable genotype: 0, 1, 2 and 3. For bi-allelic SNP sites, 0 indicates two B alleles, 1 indicates one A allele and one B allele, 2 indicates two A alleles, and 3 is a missing genotype. (When you) take out genotype data with sample and SNP IDs, and four possible values are returned 0, 1, 2 and NA (3 is replaced by NA).  

- **Colony codes alleles this way:** "The first column gives the individual ID (a string containing a maximum of 20 letters and/or numbers, no other characters are allowed), the 2nd and 3rd columns give the alleles observed for the individual at the first locus, the 4th and 5th give the alleles observed for the individual at the 2nd locus ... An allele is identified by an integer, in the range of 1~999999999. If the locus is a dominant marker, then only one (instead of 2) column is required for the marker, and the value for the genotype should be either 1 (dominant phenotype, presence of a band) or 2 (recessive phenotype, absence of a band). Missing genotypes are indicated by 0 0 for a codominant marker and 0 for a dominant marker."  Also, there cannot be any NA values in the Colony genotype data. 

I wanted to use the R program `radiator` to create a Colony formatted genotype file. But after a TON of attempts I wasn't able to get my SNPRelate .gds data into radiator directly nor could I successfully import a .vcf file. 

So, here I manually re-code the genotype matrix. 

```{r}
# Extract genotype matrix 
geno.matrix4Colony.adult <- snpgdsGetGeno(gdsobj="../qc-processing/colony/genotypes-adult-4Colony.gds", with.id=TRUE)

paste("No. of adult SNPs for Colony sibship analysis: ", ncol(geno.matrix4Colony.adult$genotype), sep="") 
paste("No. of adult samples for Colony sibship analysis: ", nrow(geno.matrix4Colony.adult$genotype), sep="") 

# Replace SNPRelate numeric allele coding with character allele coding 
a <- geno.matrix4Colony.adult$genotype %>% as_data_frame() %>% 
  mutate_all(funs(str_replace(., "0", "BB"))) %>%
  mutate_all(funs(str_replace(., "1", "AB"))) %>%
  mutate_all(funs(str_replace(., "2", "AA")))

# Duplicate each column. Now there's 2 columns per locus, both containing the same bi-allelic genotype info. 
b <- a[rep(names(a), rep(2, times=ncol(a)))] %>% clean_names()
ncol(b) == ncol(a)*2 #should be TRUE 

# For each locus, extract the first allele in column 1, and the second allele in column 2. 
c <- b  #duplicate dataframe 
odds <- seq(from = 1, to = ncol(c), by = 2)
evens <- seq(from = 2, to = ncol(c), by = 2)
odds.names <- colnames(b[odds])
evens.names <- colnames(b[evens])

for (i in 1:length(odds)) {
c[[odds.names[i]]] <- substring(b[[odds.names[i]]], first = 1, last = 1)  
}
for (i in 1:length(evens)) {
c[[evens.names[i]]] <- substring(b[[evens.names[i]]], first = 2, last = 2)  
}

# Now replace the character allele information with Colony's way of numeric coding. 
# There are 2 columns for each locus (e.g. v1 and v1_2), and only 3 options for each allele:
# 1= B allele 
# 2= A allele 
# 0= missing allele 
(geno.matrix4Colony.adult.recode <- c %>% 
  mutate_all(funs(str_replace(., "B", "1"))) %>%
  mutate_all(funs(str_replace(., "A", "2"))) %>%
  mutate(., across(everything(), ~replace_na(.x, 0))))

# Save genotype matrix to file, while also adding a column with sample number 
write_delim(cbind(sampleID=paste("O", geno.matrix4Colony.adult$sample.id, sep=""),
                  geno.matrix4Colony.adult.recode),
            file="../qc-processing/colony/geno.matrix.adult.tab", delim="   ", col_names=FALSE)
```

# Create Colony input file 

I created a text file that will be the input for Colony. I used Colony's example input file to do so. 

Here's a snapshot of the input file, truncating the genotype info (there are 53 samples, but here I only show 2)

```{bash}
head -n 28 ../qc-processing/colony/colony2_adults.dat
tail -n 12 ../qc-processing/colony/colony2_adults.dat
```

# Run Colony analysis on adult samples 

Note: I have to be in the colony directory to run it, so I change my directory in the same code chunk prior to executing the ./colony2s.out script 

```{bash, eval = FALSE}
cd ../qc-processing/colony/
./colony2s.out IFN:colony2_adults.dat
```

```{bash}
cat ../qc-processing/colony/adult_colony.BestCluster
```

```{r}
colony.clusters.adult <- read_table("../qc-processing/colony/adult_colony.BestCluster", col_names=TRUE) %>% clean_names() %>% 
  mutate_all(funs(str_replace(., "\\*|#", ""))) %>%
  mutate(cluster_index=as.numeric(cluster_index), father_id=as.numeric(father_id), mother_id=as.numeric(mother_id), probability=as.numeric(probability)) %>%
  mutate(offspring_id=str_replace(offspring_id, "O", "")) %>% as.data.frame() %>% left_join(sample.info.adult, by = c("offspring_id"="sample"))

save(colony.clusters.adult, file = "../qc-processing/colony/best-cluster-adult")  

# Plot mother ~ father ID 
#ggplotly(
  ggplot(colony.clusters.adult, aes(x=father_id, y=mother_id, size=probability, colour=population, shape=pCO2.parent)) + geom_jitter() + theme_minimal() + scale_shape_manual(values=c(1,2)) + scale_size_continuous(guide="none") #)

# How many fathers and mothers does each population + pCO2 group have? 
colony.clusters.adult %>%
  group_by(population, pCO2.parent) %>%
  summarise(n_fathers=n_distinct(father_id),n_mothers=n_distinct(mother_id)) 
  
# How many fathers and mothers does each population have
colony.clusters.adult %>%
  group_by(population) %>%
  summarise(n_fathers=n_distinct(father_id),n_mothers=n_distinct(mother_id)) 
```


## Larval Sibship analysis via Colony 

```{r}
snpset.larvae.4Colony <- snpgdsLDpruning(genofile.larvae, maf=.2, missing.rate=0.15, autosome.only=FALSE)
snpset.larvae.id.4Colony <- unlist(unname(snpset.larvae.4Colony))
snpgdsCreateGenoSet(src.fn="../qc-processing/gatk/genotypes-larvae.gds",
                    dest.fn="../qc-processing/colony/genotypes-larvae-4Colony.gds",
                    snp.id=snpset.larvae.id.4Colony)
```

## Recode genotypes for Colony 

```{r}
# Extract genotype matrix 
geno.matrix4Colony.larvae <- snpgdsGetGeno(gdsobj="../qc-processing/colony/genotypes-larvae-4Colony.gds", with.id=TRUE)

paste("No. of larvae SNPs for Colony sibship analysis: ", ncol(geno.matrix4Colony.larvae$genotype), sep="") 
paste("No. of larvae samples for Colony sibship analysis: ", nrow(geno.matrix4Colony.larvae$genotype), sep="") 

# Replace SNPRelate numeric allele coding with character allele coding 
a <- geno.matrix4Colony.larvae$genotype %>% as_data_frame() %>% 
  mutate_all(funs(str_replace(., "0", "BB"))) %>%
  mutate_all(funs(str_replace(., "1", "AB"))) %>%
  mutate_all(funs(str_replace(., "2", "AA")))

# Duplicate each column. Now there's 2 columns per locus, both containing the same bi-allelic genotype info. 
b <- a[rep(names(a), rep(2, times=ncol(a)))] %>% clean_names()
ncol(b) == ncol(a)*2 #should be TRUE 

# For each locus, extract the first allele in column 1, and the second allele in column 2. 
c <- b  #duplicate dataframe 
odds <- seq(from = 1, to = ncol(c), by = 2)
evens <- seq(from = 2, to = ncol(c), by = 2)
odds.names <- colnames(b[odds])
evens.names <- colnames(b[evens])

for (i in 1:length(odds.names)) {
c[[odds.names[i]]] <- substring(b[[odds.names[i]]], first = 1, last = 1)  
}
for (i in 1:length(evens.names)) {
c[[evens.names[i]]] <- substring(b[[evens.names[i]]], first = 2, last = 2)  
}

# Now replace the character allele information with Colony's way of numeric coding. 
# There are 2 columns for each locus (e.g. v1 and v1_2), and only 3 options for each allele:
# 1= B allele 
# 2= A allele 
# 0= missing allele 
(geno.matrix4Colony.larvae.recode <- c %>% 
  mutate_all(funs(str_replace(., "B", "1"))) %>%
  mutate_all(funs(str_replace(., "A", "2"))) %>%
  mutate(., across(everything(), ~replace_na(.x, 0))))

# Save genotype matrix to file, while also adding a column with sample number 
write_delim(cbind(sampleID=paste("O", geno.matrix4Colony.larvae$sample.id, sep=""),
                  geno.matrix4Colony.larvae.recode),
            file="../qc-processing/colony/geno.matrix.larvae.tab", delim="   ", col_names=FALSE)
```

# Create Colony input file 

I created a text file that will be the input for Colony. I used Colony's example input file as template and edited by hand. For the genotype matrix I selected all contents in the "geno.matrix.larvae.tab" file and pasted it into the text file. 

Here's a snapshot of the input file, truncating the genotype info (there are 76 samples, but here I only show 4)

```{bash}
head -n 30 ../qc-processing/colony/colony2_larvae.dat
tail -n 12 ../qc-processing/colony/colony2_larvae.dat
```

# Run Colony analysis on larvae samples 

Note: I have to be in the colony directory to run it, so I change my directory in the same code chunk prior to executing the ./colony2s.out script 

```{bash, eval = FALSE}
cd ../qc-processing/colony/
./colony2s.out IFN:colony2_larvae.dat
```

```{bash}
cat ../qc-processing/colony/larvae_colony.BestCluster
```

```{r}
(colony.clusters.larvae <- read_table("../qc-processing/colony/larvae_colony.BestCluster", col_names=TRUE) %>% clean_names() %>% 
  mutate_all(funs(str_replace(., "\\*|#", ""))) %>%
  mutate(cluster_index=as.numeric(cluster_index), father_id=as.numeric(father_id), mother_id=as.numeric(mother_id), probability=as.numeric(probability)) %>%
  mutate(offspring_id=str_replace(offspring_id, "O", "")) %>% as.data.frame() %>% left_join(sample.info.larvae, by = c("offspring_id"="sample")))

save(colony.clusters.larvae, file = "../qc-processing/colony/best-cluster-larvae")  

# Plot mother ~ father ID 
ggplotly(
  ggplot(colony.clusters.larvae, aes(x=father_id, y=mother_id, size=probability, colour=population, shape=pCO2.parent)) + geom_jitter() + theme_minimal() + scale_shape_manual(values=c(1,2)) + scale_size_continuous(guide="none"))

# How many fathers and mothers does each population + pCO2 group have? 
colony.clusters.larvae %>%
  group_by(population, pCO2.parent) %>%
  summarise(n_fathers=n_distinct(father_id),n_mothers=n_distinct(mother_id)) %>%
  pivot_longer(names_to = "parent", values_to="n_parents", cols = c(n_fathers, n_mothers)) %>% 
  mutate(population=as.factor(population),pCO2.parent=as.factor(pCO2.parent),parent=as.factor(parent)) %>%
  ggplot(aes(fill=population, alpha=pCO2.parent, x=population, y=n_parents)) + geom_bar(position="dodge", stat="identity") +
  scale_alpha_discrete(range = c(0.35, 0.8)) + ylab("No. of parents") + 
  facet_wrap(~parent) + theme_minimal() + theme(axis.text.x = element_blank()) + xlab("Population & parental pCO2 exposure")
```

## Juvenile Sibship analysis via Colony 

```{r}
snpset.juvenile.4Colony <- snpgdsLDpruning(genofile.juvenile, maf=.4, missing.rate=0, autosome.only=FALSE)
snpset.juvenile.id.4Colony <- unlist(unname(snpset.juvenile.4Colony))
snpgdsCreateGenoSet(src.fn="../qc-processing/gatk/genotypes-juvenile.gds",
                    dest.fn="../qc-processing/colony/genotypes-juvenile-4Colony.gds",
                    snp.id=snpset.juvenile.id.4Colony)
```

## Recode genotypes for Colony 

```{r}
# Extract genotype matrix 
geno.matrix4Colony.juvenile <- snpgdsGetGeno(gdsobj="../qc-processing/colony/genotypes-juvenile-4Colony.gds", with.id=TRUE)

paste("No. of juvenile SNPs for Colony sibship analysis: ", ncol(geno.matrix4Colony.juvenile$genotype), sep="") 
paste("No. of juvenile samples for Colony sibship analysis: ", nrow(geno.matrix4Colony.juvenile$genotype), sep="") 

# Replace SNPRelate numeric allele coding with character allele coding 
a <- geno.matrix4Colony.juvenile$genotype %>% as_data_frame() %>% 
  mutate_all(funs(str_replace(., "0", "BB"))) %>%
  mutate_all(funs(str_replace(., "1", "AB"))) %>%
  mutate_all(funs(str_replace(., "2", "AA")))

# Duplicate each column. Now there's 2 columns per locus, both containing the same bi-allelic genotype info. 
b <- a[rep(names(a), rep(2, times=ncol(a)))] %>% clean_names()
ncol(b) == ncol(a)*2 #should be TRUE 

# For each locus, extract the first allele in column 1, and the second allele in column 2. 
c <- b  #duplicate dataframe 
odds <- seq(from = 1, to = ncol(c), by = 2)
evens <- seq(from = 2, to = ncol(c), by = 2)
odds.names <- colnames(b[odds])
evens.names <- colnames(b[evens])

for (i in 1:length(odds.names)) {
c[[odds.names[i]]] <- substring(b[[odds.names[i]]], first = 1, last = 1)  
}
for (i in 1:length(evens.names)) {
c[[evens.names[i]]] <- substring(b[[evens.names[i]]], first = 2, last = 2)  
}

# Now replace the character allele information with Colony's way of numeric coding. 
# There are 2 columns for each locus (e.g. v1 and v1_2), and only 3 options for each allele:
# 1= B allele 
# 2= A allele 
# 0= missing allele 
(geno.matrix4Colony.juvenile.recode <- c %>% 
  mutate_all(funs(str_replace(., "B", "1"))) %>%
  mutate_all(funs(str_replace(., "A", "2"))) %>%
  mutate(., across(everything(), ~replace_na(.x, 0))))

# Save genotype matrix to file, while also adding a column with sample number 
write_delim(cbind(sampleID=paste("O", geno.matrix4Colony.juvenile$sample.id, sep=""),
                  geno.matrix4Colony.juvenile.recode),
            file="../qc-processing/colony/geno.matrix.juvenile.tab", delim="   ", col_names=FALSE)
```

# Create Colony input file 

I created a text file that will be the input for Colony. I used Colony's example input file as template and edited by hand. For the genotype matrix I selected all contents in the "geno.matrix.juvenile.tab" file and pasted it into the text file. 

Here's a snapshot of the input file, truncating the genotype info (there are 16 samples, but here I only show 1)

```{bash}
head -n 28 ../qc-processing/colony/colony2_juveniles.dat
tail -n 12 ../qc-processing/colony/colony2_juveniles.dat
```

# Run Colony analysis on juvenile samples 

Note: I have to be in the colony directory to run it, so I change my directory in the same code chunk prior to executing the ./colony2s.out script 

```{bash, eval = FALSE}
cd ../qc-processing/colony/
./colony2s.out IFN:colony2_juveniles.dat
```

```{bash}
cat ../qc-processing/colony/juvenile_colony.BestCluster
```

```{r}
(colony.clusters.juvenile <- read_table("../qc-processing/colony/juvenile_colony.BestCluster", col_names=TRUE) %>% clean_names() %>% 
  mutate_all(funs(str_replace(., "\\*|#", ""))) %>%
  mutate(cluster_index=as.numeric(cluster_index), father_id=as.numeric(father_id), mother_id=as.numeric(mother_id), probability=as.numeric(probability)) %>%
  mutate(offspring_id=str_replace(offspring_id, "O", "")) %>% as.data.frame() %>% left_join(sample.info.juvenile, by = c("offspring_id"="sample")))

save(colony.clusters.juvenile, file = "../qc-processing/colony/best-cluster-juvenile")  

# Plot mother ~ father ID 
ggplotly(
  ggplot(colony.clusters.juvenile, aes(x=father_id, y=mother_id, size=probability, colour=population, shape=pCO2.parent)) + geom_jitter() + theme_minimal() + scale_shape_manual(values=c(1,2)) + scale_size_continuous(guide="none"))

# How many fathers and mothers does each population + pCO2 group have? 
colony.clusters.juvenile %>%
  group_by(population, pCO2.parent) %>%
  summarise(n_fathers=n_distinct(father_id),n_mothers=n_distinct(mother_id)) %>%
  pivot_longer(names_to = "parent", values_to="n_parents", cols = c(n_fathers, n_mothers)) %>% 
  mutate(population=as.factor(population),pCO2.parent=as.factor(pCO2.parent),parent=as.factor(parent)) %>%
  ggplot(aes(fill=population, alpha=pCO2.parent, x=population, y=n_parents)) + geom_bar(position="dodge", stat="identity") +
  scale_alpha_discrete(range = c(0.35, 0.8)) + ylab("No. of parents") + 
  facet_wrap(~parent) + theme_minimal() + theme(axis.text.x = element_blank()) + xlab("Population & parental pCO2 exposure")
```
