---
title: "QuantSeq-2020-Importing-STAR"
author: "Laura H Spencer"
date: "5/18/2020"
output: html_document
---

### Load libraries

```{r message=FALSE, warning=FALSE, results=FALSE}
list.of.packages <- c("tidyverse", "reshape2", "here", "plotly", "purrr") #add new libraries here 

new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

# Load all libraries 
lapply(list.of.packages, FUN = function(X) {
  do.call("require", list(X)) 
})
```



```{r}
sample.info <- read.csv("../raw-data/quantseq2020_key.csv", header=T, na.strings = NA) %>%
          mutate_at(vars(lane, temp.parent), as.factor) %>% 
          mutate_at(vars(RNA.conc, endpoint.RFU, bioan.mean.bp), as.numeric) %>%
          mutate(sample_stage=paste(sample, stage, sep="_"))

filenames <- data.frame(read.table("../qc-processing/STAR/test-out/countsfilenames.txt", header = F, stringsAsFactors = F, fill = FALSE)) %>%
          dplyr::rename(filename = 1, sample=2) %>% 
          mutate_at(vars(sample), as.factor) %>% 
          left_join(sample.info[c("sample", "sample_stage")])

files <- file.path(filenames$filename) #extract vector of filenames 
all(file.exists(files)) #easy code to check that all files exist! 
names(files) <- filenames$sample_stage #name each filename with sample#_type
key <- full_join(sample.info, filenames) 

# save(key, file="../raw-data/key")
```

## Read in count files, selecting only the first 2 columns (1=gene ID, 2=count); rename columns 

```{r}
file_list <- vector(mode = "list", length = nrow(filenames))
names(file_list) <- c(filenames$sample_stage)

for (i in 1:nrow(filenames)) {
    file_list[[i]] <- data.frame(read.delim(file=files[i]))[-1:-4,1:2]
    names(file_list[[i]]) <- c("gene", filenames$sample_stage[i])
    print(paste("Total COUNTS,", names(file_list[[i]][2]), ":", prettyNum(sum(file_list[[i]][2]), big.mark=","), sep=" "))
    print(paste("Total GENES,", names(file_list[[i]][2]), ":", prettyNum(nrow(file_list[[i]] %>% filter(.[[2]] != 0)), big.mark=","), sep=" "))
}
```

```{r}
counts <- file_list %>% purrr::reduce(full_join, by = "gene") %>% column_to_rownames(var="gene") 
```

### Summarize counts across all samples  

```{r}
print(paste("Number of samples:", ncol(counts), sep=" "))
print(paste("Total number of genes in dataframe (before filtering):", prettyNum(nrow(counts), big.mark = ","), sep=" "))
print(paste("Total counts, all samples:", prettyNum(sum(counts), big.mark = ","), sep=" "))
```

### Transpose dataframe so each row = a sample (aka "objects"), and each column = genes (aka "variables") 
```{r}
counts.t <- t(counts) #transform data to have each sample a row, each column a gene 
```

### Remove genes that don't show up in any samples (why are they here?)
```{r}
source("../references/biostats.R")

100*(1/nrow(counts.t)) #if just one sample had a gene, what would that % be? 12.5. Use below to drop. 
counts.ts <- as.data.frame(drop.var(counts.t, min.po = ceiling(100*(1/nrow(counts.t)))))
print(paste("# of genes not present in samples & dropped:", ncol(counts.t) - ncol(counts.ts), sep=" "))
print(paste("# of genes remaimning:", ncol(counts.ts), sep=" "))
```

## Drop genes with very little variability between all samples, using minimum coefficient variation (min.cv)=5
```{r}
counts.tsv <- as.data.frame(drop.var(counts.ts, min.cv=5)) 
print(paste("number of genes dropped due to low variance", ncol(counts.ts) - ncol(counts.tsv)))
print(paste("# number of genes remain in dataset after filtering: ", ncol(counts.tsv))) 
```

```{r}
for (i in (1:nrow(counts.tsv))) {
  print(paste("Total GENES for", rownames(counts.tsv)[i], ":", prettyNum(ncol(counts.tsv[i,counts.tsv[i,]!=0]), big.mark=","), sep=" "))
}

print(paste("Total COUNTS for", rownames(counts.tsv), ":",  prettyNum(rowSums(counts.ts), big.mark = ","), sep=" "))
barplot(rowSums(counts.tsv), main="Total counts by sample", las=2) #inspect total counts by sample

#save(counts.ts, file = "../results/counts.ts")
```

